---
title: "Study1_HRV_ENreg"
author: "Allison Eriksson"
date: "`r Sys.Date()`"
output:
  html_document
    toc: true
    toc_float: true
    collapsed: false
    number_sections: true
    editor_options:
  chunk_output_type: console
---
Install packages
```{r setup, include=FALSE}

library(naniar) #for checking data

library(glmnet)

library(foreign)

library(pROC)

library(doParallel)

library(tidyr)

library(ggplot2)

library(mice) #imputation

library(tidyverse)

library(MASS)


library(lars)

library(caret) #LOOCV 

library(dplyr)

library(generalCorr)

library(Hmisc)

library(lattice)

pkgs <- list("glmnet", "doParallel", "foreach","ggplot2", "pROC", "naniar", "mice", "tidyverse", "lars", "caret", "generalCorr", "MASS", "dplyr", "Hmisc", "lattice")
lapply(pkgs, require, character.only = T)
registerDoParallel(cores = 5)

```
# DATA SETUP
## Check data
```{r setup, include=FALSE}
#from website https://www.r-bloggers.com/2017/09/variable-selection-with-elastic-net/ 
 

hist(HRV_HFpow1)
 hist(HRV_HFpow2)
 hist(HRV_LFHF_1) #skewed 
 hist(HRV_LFHF_2)#skewed 
 hist(HRV_LFpow_1)
 hist(HRV_LFpow_2)
 hist(HRV_POW1)
 hist(HRV_POW2)
 hist(HRV_RMSSD_1)#skewed 
 hist(HRV_RMSSD_2)
 hist(HRV_SDNN_1)
 hist(HRV_SDNN_2)
 hist(HRV_TP_1)
 hist(HRV_TP_2)
 hist(scale(HRV_LFHF_2))
hist(HRV_LFHF_2)


```

## VARIABLE TRANSFORMATION
```{r }
##https://towardsdatascience.com/smart-handling-of-missing-data-in-r-6425f8a559f2

### Transform variables to df1, add original and transformed variables to df_short list 
#imported spss file manually
df1 <- BASIC_updated_22nov_na_rm
#df1 = read.spss("/Users/aller405/Desktop/BASIC_updated_22nov_na.rm.sav", to.data.frame=TRUE,)

POB <- as.numeric(df1$v17_fodelseort_R)
df1$POB <- POB

# POBtext <- (df1$v17_fodelseort_R)
# df1$POBtext <- POBtext

pp_EPDS <- as.numeric(df1$ppv6_EPDS_10_11R)
df1$pp_EPDS <- pp_EPDS

ppEPDStot <- as.numeric(df1$ppv6_EPDS_9R)
df1$ppEPDStot <- ppEPDStot

pp_BECK <- as.numeric(df1$ppv6_beck_score_R_01)
df1$pp_BECK <- pp_BECK

ppBECKtot <- as.numeric(df1$ppv6_beck_tot_R)
df1$ppBECKtot <- ppBECKtot


EDU <- as.numeric(as.numeric(df1$v17_utbildning_R_01))
df1$EDU <- EDU

# EDUtext <- (as.numeric(df1$v17_utbildning_R_01))
# df1$EDUtext <- EDUtext

SLEEP <- as.numeric(as.numeric(df1$v32_somn))
df1$SLEEP <- SLEEP

# SLEEPtext <- (as.numeric(df1$v32_somn))
# df1$SLEEPtext <- SLEEPtext

AGE <- as.numeric(df1$B_alder_groups)
df1$AGE <- AGE 

AGE_tot <- (as.numeric(df1$NK_Age_at_partus))
df1$AGE_tot <- AGE_tot 

EPDS32 <- as.numeric(df1$v32_EPDS_D_9R)
df1$EPDS32 <- EPDS32

EPDS_32_tot <- as.numeric(df1$v32_EPDS_9R)
df1$EPDS_32_tot <- EPDS_32_tot

BECK_32_tot <- as.numeric(df1$v32_beck_tot_R)
df1$BECK_32_tot <- BECK_32_tot

BECK32 <- as.numeric(df1$v32_beck_score_R_01)
df1$BECK32 <- BECK32

FOD <- as.numeric(df1$NK_delivery_fear_2)
df1$FOD <- FOD

# FODtext <- (df1$NK_delivery_fear_2)
# df1$FODtext <- FODtext

# #### Figure out how to dichotomize this variable- less than 65 is low resilience. Try to transform in short df after imputation? 
# df1$RS_14 <- factor(df1$v32_RS_14_SCORE_R)
# df1$RS_14[df1$v32_RS_14_SCORE_R =="Very low"] <- "0"
# df1$RS_14[df1$v32_RS_14_SCORE_R =="Low"] <- "0"
# df1$RS_14[df1$v32_RS_14_SCORE_R =="On the low end"] <- "0"
# df1$RS_14[df1$v32_RS_14_SCORE_R =="On the low end"] <- "0"
# df1$RS_14[df1$v32_RS_14_SCORE_R =="Moderate"] <- "1"
# df1s$RS_14[df1$v32_RS_14_SCORE_R =="Moderately high"] <- "1"
# df1$RS_14[df1$v32_RS_14_SCORE_R =="High"] <- "1"

RS_14_tot <- as.numeric(df1$v32_RS_14_R)
df1$RS_14_tot <- RS_14_tot


BMI_tot <- as.numeric(df1$v32_BMI)
df1$BMI_tot <- BMI_tot

BMI <- as.numeric(df1$v32_BMI_groups_R)
df1$BMI <- BMI

PAR <- as.numeric(df1$NK_parity_D)
df1$PAR <- PAR

# PARtext<- (df1$NK_parity)
# df1$PARtext <- PARtext

ILL <- as.numeric(df1$Illness_combined)
df1$ILL <- ILL

# ILLtext <- (df1$Illness_combined_variables)
# df1$ILLtext <- ILLtext

DEPH <- as.numeric(df1$NK_depression_history)
df1$DEPH <- DEPH

# DEPHtext <- (df1$NK_depression_history)
# df1$DEPHtext <- DEPHtext

SSRI <- as.numeric(df1$NK_SSRI_pregnancy_loose)
df1$SSRI <- SSRI

# SSRItext <- (df1$NK_SSRI_pregnancy_loose)
# df1$SSRItext <- SSRItext

df1$NK_SSRI_pregnancy_loose

HRV_HFpow_1<- df1$HRV_HFpower_pregnv1
df1$HRV_HFpow_1 <- HRV_HFpow_1

HRV_HFpow_2 <- (df1$HRV_HFpower_pregnv2)
df1$HRV_HFpow_2 <- HRV_HFpow_2


HRV_LFpow_1 <- (df1$HRV_LFpower_pregnv1)
df1$HRV_LFpow_1 <- HRV_LFpow_1

HRV_LFpow_2 <- (df1$HRV_LFpower_pregnv2)
df1$HRV_LFpow_2 <- HRV_LFpow_2

HRV_TP_1 <- (df1$HRV_Totalpower_pregnv1)
df1$HRV_TP_1 <- HRV_TP_1

HRV_TP_2 <- (df1$HRV_Totalpower_pregnv2)
df1$HRV_TP_2 <- HRV_TP_2


HRV_RMSSD_1 <- (df1$HRV_RMSSD_pregnv1)
df1$HRV_RMSSD_1 <- HRV_RMSSD_1

HRV_RMSSD_2 <- (df1$HRV_RMSSD_pregnv2)
df1$HRV_RMSSD_2 <- HRV_RMSSD_2

HRV_SDNN_1 <- (df1$HRV_SDNNms_pregnv1)
df1$HRV_SDNN_1 <- HRV_SDNN_1

HRV_SDNN_2 <- (df1$HRV_SDNNms_pregnv2)
df1$HRV_SDNN_2 <- HRV_SDNN_2

HRV_LFHF_1 <- (df1$HRV_LFHFratio_pregnv1)
df1$HRV_LFHF_1 <- HRV_LFHF_1

HRV_LFHF_2 <- (df1$HRV_LFHFratio_pregnv2)
df1$HRV_LFHF_2 <- HRV_LFHF_2


```

##Select variables of interest

```{r }
# if error in select: https://www.geeksforgeeks.org/how-to-fix-error-in-select-unused-arguments-in-r/
#make a selection without outcome variables for imputation (cannot impute outcome varaibles)! 
df1_tib <- as_tibble(df1)

df_short_pred_only <- df1_tib %>% dplyr::select(POB, EDU, SLEEP, AGE_tot, BMI_tot, BMI, PAR, ILL, DEPH, SSRI, FOD, RS_14_tot, 
                                                EPDS_32_tot, BECK_32_tot , HRV_HFpow_1, HRV_HFpow_2, HRV_LFpow_1, HRV_LFpow_2, HRV_TP_1, HRV_TP_2,
                                                HRV_RMSSD_1, HRV_RMSSD_2, HRV_SDNN_1, HRV_SDNN_2, HRV_LFHF_1, HRV_LFHF_2)    


#Check correlations

cormatrix <- rcorr(as.matrix(df_short_pred_only, type = "pearson"))
cormatrix
cormatrixP <- round(cormatrix$P,2)
cormatrixN <- round(cormatrix$n,2)
cormatrixR <- round(cormatrix$r,2)

gc()
```
# IMPUTATION
## Check data before imputation
```{r}
# Are there missing values in the dataset?
any_na(df_short_pred_only)
# How many?
n_miss(df_short_pred_only)
prop_miss(df_short_pred_only)
# Which variables are affected?
df_short_pred_only %>% is.na() %>% colSums()

# Get number of missings per variable (n and %)
miss_var_summary(df_short_pred_only)
miss_var_table(df_short_pred_only)
# Get number of missings per participant (n and %)
miss_case_summary(df_short_pred_only)
miss_case_table(df_short_pred_only)

# Which variables contain the most missing variables?
gg_miss_var(df_short_pred_only)

# Where are missings located?
vis_miss(df_short_pred_only) + theme(axis.text.x = element_text(angle=80))

# Which combinations of variables occur to be missing together?
gg_miss_upset(df_short_pred_only)


# # Get a heatmap of missingness broken down by ppBECK
# is.factor(df_short$ppBECK)
# gg_miss_fct(df_short, fct = ppBECK)



```

## MICE Imputation

```{r}

tbl_df_short_pred <- as_tibble(df_short_pred_only) #### Why should it be in tbl format? 
df_short_pred_mat <- quickpred(tbl_df_short_pred, mincor = 0.25)
df_short_imp_pred <- mice(tbl_df_short_pred, m = 10, meth='pmm', seed = 14221, predictorMatrix = df_short_pred_mat)
df_pred_complete <- complete(df_short_imp_pred)

any_na(df_pred_complete)
gc()


```

# EPDS SETUP
```{r}
# Create a dataset with ppEPDS variable + predictors 
df_epds_pred_complete <- df_pred_complete
df_epds_pred_complete$pp_EPDS <- pp_EPDS
df_epds_pred_complete <- df_epds_pred_complete[!is.na(df_epds_pred_complete$pp_EPDS),]

#Create factors with levels 
df_epds_pred_complete$pp_EPDS <- 
  factor(df_epds_pred_complete$pp_EPDS, 
         levels=c(0,1),
         labels=c("Non-PPD", # Reference
                  "PPD"))
```

# BECK SETUP
```{r}
 #####
### Adding BAI to imputed variable data 
gc()
# Create a dataset with ppBECK variable + predictors 
df_beck_pred_complete <- df_pred_complete
df_beck_pred_complete$pp_BECK <- pp_BECK
df_beck_pred_complete <- df_beck_pred_complete[!is.na(df_beck_pred_complete$pp_BECK),]
#df_beck_pred_complete$pp_BECK <- recode(df_beck_pred_complete$pp_BECK, 
                                        #`1` = "0", `2` = "1")
#Create factors with levels 
df_beck_pred_complete$pp_BECK <- 
  factor(df_beck_pred_complete$pp_BECK, 
         levels=c(0,1))
        # labels=c("Non-anxious", # Reference
                 # "Anxious"))
```


```{r}

```

# TABLE 2

```{r}
# Dep histogram
histogram(df_epds_pred_complete$HRV_HFpow_1)
histogram(df_epds_pred_complete$HRV_HFpow_2)
histogram(df_epds_pred_complete$HRV_LFpow_1)
histogram(df_epds_pred_complete$HRV_LFpow_2)
histogram(df_epds_pred_complete$HRV_TP_1)
histogram(df_epds_pred_complete$HRV_TP_2)
histogram(df_epds_pred_complete$HRV_LFHF_1)
histogram(df_epds_pred_complete$HRV_LFHF_2)
histogram(df_epds_pred_complete$HRV_RMSSD_1) # Skewed
histogram(df_epds_pred_complete$HRV_RMSSD_2) # Skewed
histogram(df_epds_pred_complete$HRV_SDNN_1)
histogram(df_epds_pred_complete$HRV_SDNN_2)

# HRV/DEP group differences 
t.test(df_epds_pred_complete$HRV_HFpow_1  ~  df_epds_pred_complete$pp_EPDS)

t.test(df_epds_pred_complete$HRV_HFpow_2  ~  df_epds_pred_complete$pp_EPDS)

t.test(df_epds_pred_complete$HRV_LFpow_1 ~ df_epds_pred_complete$pp_EPDS)

t.test(df_epds_pred_complete$HRV_LFpow_2 ~ df_epds_pred_complete$pp_EPDS)

t.test(df_epds_pred_complete$HRV_TP_1 ~ df_epds_pred_complete$pp_EPDS)

t.test(df_epds_pred_complete$HRV_TP_2 ~ df_epds_pred_complete$pp_EPDS)

df_epds_pred_complete$pp_EPDS <- as.numeric(df_epds_pred_complete$pp_EPDS)
wilcox.test(df_epds_pred_complete$HRV_RMSSD_1, df_epds_pred_complete$pp_EPDS)

wilcox.test(df_epds_pred_complete$HRV_RMSSD_2, df_epds_pred_complete$pp_EPDS)

t.test(df_epds_pred_complete$HRV_SDNN_1 ~ df_epds_pred_complete$pp_EPDS)

t.test(df_epds_pred_complete$HRV_SDNN_2 ~ df_epds_pred_complete$pp_EPDS)


# ANX histogram

histogram(df_beck_pred_complete$HRV_HFpow_1)
histogram(df_beck_pred_complete$HRV_HFpow_2)
histogram(df_beck_pred_complete$HRV_LFpow_1)
histogram(df_beck_pred_complete$HRV_LFpow_2)
histogram(df_beck_pred_complete$HRV_TP_1)
histogram(df_beck_pred_complete$HRV_TP_2)
histogram(df_beck_pred_complete$HRV_LFHFpow_1)
histogram(df_beck_pred_complete$HRV_LFHFpow_2)
histogram(df_beck_pred_complete$HRV_RMSSD_1) # Skewed
histogram(df_beck_pred_complete$HRV_RMSSD_2) # Skewed
histogram(df_beck_pred_complete$HRV_SDNN_1)
histogram(df_beck_pred_complete$HRV_SDNN_2)

#ANX/HRV group differences
t.test(df_beck_pred_complete$HRV_HFpow_1  ~  df_beck_pred_complete$pp_BECK)

t.test(df_beck_pred_complete$HRV_HFpow_2  ~  df_beck_pred_complete$pp_BECK)

t.test(df_beck_pred_complete$HRV_LFpow_1  ~  df_beck_pred_complete$pp_BECK)

t.test(df_beck_pred_complete$HRV_LFpow_2  ~  df_beck_pred_complete$pp_BECK)

t.test(df_beck_pred_complete$HRV_TP_1  ~  df_beck_pred_complete$pp_BECK)

t.test(df_beck_pred_complete$HRV_TP_2  ~  df_beck_pred_complete$pp_BECK)

df_beck_pred_complete$pp_BECK <- as.numeric(df_beck_pred_complete$pp_BECK)

wilcox.test(df_beck_pred_complete$HRV_LFHF_1  ~  df_beck_pred_complete$pp_BECK)

wilcox.test(df_beck_pred_complete$HRV_LFHF_2  ~  df_beck_pred_complete$pp_BECK)

wilcox.test(df_beck_pred_complete$HRV_RMSSD_1  ~  df_beck_pred_complete$pp_BECK)


wilcox.test(df_beck_pred_complete$HRV_RMSSD_2  ~  df_beck_pred_complete$pp_BECK)

t.test(df_beck_pred_complete$HRV_SDNN_1  ~  df_beck_pred_complete$pp_BECK)

t.test(df_beck_pred_complete$HRV_SDNN_2  ~  df_beck_pred_complete$pp_BECK)

```




```{r}


```


## ELASTIC NET REGRESSION with DEP outcome for whole dataset with all predictors imputed
```{r}
set.seed(123987)  #123987 used in final model
n <- nrow(df_epds_pred_complete)
sample <- sample(seq(n), size = n * 0.8, replace = FALSE) 
train <- df_epds_pred_complete[ sample, ]
test <- df_epds_pred_complete[-sample,]
# str(train)
# summary(train)
# names(df_epds_pred_complete)
# any_na(df_epds_pred_complete)
# class(df_epds_pred_complete$pp_EPDS)


mdlY <- as.factor(data.matrix(train[,"pp_EPDS" ],))
mdlX <-scale (as.matrix(train[,c("POB", "EDU", "SLEEP", "AGE_tot", "BMI_tot","PAR", "ILL", "DEPH", "SSRI", "FOD", "RS_14_tot",
                                 "EPDS_32_tot", "BECK_32_tot" , "HRV_HFpow_1", "HRV_HFpow_2" , "HRV_LFpow_1" , "HRV_LFpow_2", "HRV_TP_1", "HRV_TP_2",
                                 "HRV_RMSSD_1", "HRV_RMSSD_2", "HRV_SDNN_1", "HRV_SDNN_2","HRV_LFHF_1", "HRV_LFHF_2"), ]))
newY <- as.factor(data.matrix(test[,"pp_EPDS"], ))
newX <- scale (as.matrix(test[,c("POB", "EDU", "SLEEP", "AGE_tot", "BMI_tot","PAR", "ILL", "DEPH", "SSRI", "FOD", "RS_14_tot",
                                 "EPDS_32_tot", "BECK_32_tot" , "HRV_HFpow_1", "HRV_HFpow_2" , "HRV_LFpow_1" , "HRV_LFpow_2", "HRV_TP_1", "HRV_TP_2",
                                 "HRV_RMSSD_1", "HRV_RMSSD_2", "HRV_SDNN_1", "HRV_SDNN_2","HRV_LFHF_1", "HRV_LFHF_2"), ]))

### removed "RS_14_tot",
#### Elastic net 
a <- seq(0.1, 0.9, 0.05)
set.seed(123987)
search <- foreach(i = a, .combine = rbind) %dopar% {
  cv <- cv.glmnet(mdlX, mdlY, family = "binomial", nfold = 10, type.measure = "deviance", parallel  = TRUE, alpha = i)
  data.frame(cvm = cv$cvm[cv$lambda == cv$lambda.1se], lambda.1se = cv$lambda.1se, alpha = i)
}
cv3 <- search[search$cvm == min(search$cvm), ]
md_dep <- glmnet(mdlX, mdlY, family = "binomial", lambda = cv3$lambda.1se, alpha = cv3$alpha)
#md_dep <- glmnet(mdlX, mdlY, family = "binomial", lambda = .25, alpha = .25)
coef(md_dep)

roc(newY, as.numeric(predict(md_dep, newX, type = "response")))


ROC_Final_EPDS_Com = roc(newY, as.numeric(predict(md_dep, newX, type = "response")))

ggroc(ROC_Final_EPDS_Com, colour = 'steelblue', legacy.axes = TRUE, size = 2) + 
  theme(text=element_text(size= 16)) +
  ggtitle(paste0('Elastic net regression with depression outcome ',
                 '(AUC = ', round(ROC_Final_EPDS_Com$auc,4),')'))

```
## Output
```{r}
roc(newY, as.numeric(predict(md_dep, newX, type = "response")))


ROC_Final_EPDS_Com = roc(newY, as.numeric(predict(md_dep, newX, type = "response")))

ggroc(ROC_Final_EPDS_Com, colour = 'steelblue', legacy.axes = TRUE, size = 2) + 
  theme(text=element_text(size= 16)) +
  ggtitle(paste0('Elastic net regression with depression outcome ',
                 '(AUC = ', round(ROC_Final_EPDS_Com$auc,4),')'))


# 26 x 1 sparse Matrix of class "dgCMatrix"
# s0
# (Intercept) -1.13427003
# POB          .         
# EDU          .         
# SLEEP        .         
# AGE_tot      .         
# BMI          .         
# PAR          .         
# ILL          .         
# DEPH         .         
# SSRI         .         
# FOD          .         
# RS_14_tot   -0.23200202
# EPDS_32_tot  0.26697753
# BECK_32_tot  0.14608420
# HRV_HFpow_1 -0.08520759
# HRV_HFpow_2  .         
# HRV_LFpow_1 -0.03675938
# HRV_LFpow_2  .         
# HRV_TP_1    -0.03835127
# HRV_TP_2    -0.04113627
# HRV_RMSSD_1  .         
# HRV_RMSSD_2  .         
# HRV_SDNN_1   .         
# HRV_SDNN_2  -0.04582977
# HRV_LFHF_1   .         
# HRV_LFHF_2   .         
# > roc(newY, as.numeric(predict(md_dep, newX, type = "response")))
# Setting levels: control = Non-PPD, case = PPD
# Setting direction: controls < cases
# 
# Call:
#   roc.default(response = newY, predictor = as.numeric(predict(md_dep,     newX, type = "response")))
# 
# Data: as.numeric(predict(md_dep, newX, type = "response")) in 20 controls (newY Non-PPD) < 3 cases (newY PPD).
# Area under the curve: 0.9333

#lambda = 0.2500912
#alpha = 0.25

# Setting alpha = .25 as in final model
# cv1 <- cv.glmnet(mdlX, mdlY, family = "binomial", nfold = 10,type.measure = "deviance", parallel = TRUE, alpha = .25)
# md1 <- glmnet(mdlX, mdlY, family = "binomial", lambda = .25, alpha = .25)
# coef(md1)
# 
# roc(newY, as.numeric(predict(md1, newX, type = "response")))



```


#BECK SETUP (rerun before doing EN!!)
```{r}
 #####
### Adding BAI to imputed variable data 
gc()
# Create a dataset with ppBECK variable + predictors 
df_beck_pred_complete <- df_pred_complete
df_beck_pred_complete$pp_BECK <- pp_BECK
df_beck_pred_complete <- df_beck_pred_complete[!is.na(df_beck_pred_complete$pp_BECK),]
#df_beck_pred_complete$pp_BECK <- recode(df_beck_pred_complete$pp_BECK, 
                                        #`1` = "0", `2` = "1")
#Create factors with levels 
df_beck_pred_complete$pp_BECK <- 
  factor(df_beck_pred_complete$pp_BECK, 
         levels=c(0,1))
        # labels=c("Non-anxious", # Reference
                 # "Anxious"))

```



## EN reg for ANX outcome with all predictors 
```{r}
set.seed(123987)
n <- nrow(df_beck_pred_complete)
sample <- sample(seq(n), size = n * 0.5, replace = FALSE) 
train <- df_beck_pred_complete[sample,]
test <- df_beck_pred_complete[-sample,]
class(df_beck_pred_complete$pp_BECK)
# names(df_beck_pred_complete)
# any_na(df_beck_pred_complete)

mdlY <- (data.matrix(train[,"pp_BECK" ],))

mdlX <-scale(data.matrix(train[,c( "POB", "EDU", "SLEEP", "AGE_tot", "BMI_tot","PAR", "ILL", "DEPH", "SSRI", "FOD", "RS_14_tot", 
                                   "EPDS_32_tot", "BECK_32_tot" , "HRV_HFpow_1", "HRV_HFpow_2" , "HRV_LFpow_1" , "HRV_LFpow_2", "HRV_TP_1", "HRV_TP_2",
                                   "HRV_RMSSD_1", "HRV_RMSSD_2", "HRV_SDNN_1", "HRV_SDNN_2","HRV_LFHF_1", "HRV_LFHF_2"), ]))
newY <- (data.matrix(test[,"pp_BECK"], ))
newX <- scale(data.matrix(test[,c("POB", "EDU", "SLEEP", "AGE_tot", "BMI_tot","PAR", "ILL", "DEPH", "SSRI", "FOD", "RS_14_tot", 
                                  "EPDS_32_tot", "BECK_32_tot" , "HRV_HFpow_1", "HRV_HFpow_2" , "HRV_LFpow_1" , "HRV_LFpow_2", "HRV_TP_1", "HRV_TP_2",
                                  "HRV_RMSSD_1", "HRV_RMSSD_2", "HRV_SDNN_1", "HRV_SDNN_2","HRV_LFHF_1", "HRV_LFHF_2"), ]))



a <- seq(0.1, 0.9, 0.05)
set.seed(123987)
search <- foreach(i = a, .combine = rbind) %dopar% {
  cv <- cv.glmnet(mdlX, mdlY, family = "binomial", nfold = 10, type.measure = "deviance", parallel  = TRUE, alpha = i)
  data.frame(cvm = cv$cvm[cv$lambda == cv$lambda.1se], lambda.1se = cv$lambda.1se, alpha = i)
}
cv3 <- search[search$cvm == min(search$cvm), ]
md_anx <- glmnet(mdlX, mdlY, family = "binomial", lambda = .15, alpha = .35)
#md_anx <- glmnet(mdlX, mdlY, family = "binomial", lambda = cv3$lambda.1se, alpha = cv3$alpha)
coef(md_anx)
#class(newY)
roc(as.factor(newY), predict(md_anx, newX, type = "response", direction = ">"))
ROC_EN_BECK_Com = roc(as.factor(newY), predict(md_anx, newX, type = "response", direction = ">"))
as.factor(as.numeric(newY))

ggroc(ROC_EN_BECK_Com, colour = 'steelblue', legacy.axes = TRUE, size = 2) + 
  theme(text=element_text(size= 16)) +
  ggtitle(paste0('Elastic net regression with anxiety outcome ',
                 '(AUC = ', round(ROC_EN_BECK_Com$auc,4),')'))
```
# SUPP MODELS     
## EN Reg for DEP T1 only 
```{r}
set.seed(123987)  #123987 used in final model
n <- nrow(df_epds_pred_complete)
sample <- sample(seq(n), size = n * 0.8, replace = FALSE) 
train <- df_epds_pred_complete[ sample, ]
test <- df_epds_pred_complete[-sample,]
# str(train)
# summary(train)
# names(df_epds_pred_complete)
# any_na(df_epds_pred_complete)
# class(df_epds_pred_complete$pp_EPDS)


mdlY <- as.factor(data.matrix(train[,"pp_EPDS" ],))
mdlX <-scale (as.matrix(train[,c("POB", "EDU", "SLEEP", "AGE_tot", "BMI_tot","PAR", "ILL", "DEPH", "SSRI", "FOD", "RS_14_tot",
                                 "EPDS_32_tot", "BECK_32_tot" ), ]))
newY <- as.factor(data.matrix(test[,"pp_EPDS"], ))
newX <- scale (as.matrix(test[,c("POB", "EDU", "SLEEP", "AGE_tot", "BMI_tot","PAR", "ILL", "DEPH", "SSRI", "FOD", "RS_14_tot",
                                 "EPDS_32_tot", "BECK_32_tot"), ]))

### removed "RS_14_tot",
#### Elastic net 
a <- seq(0.1, 0.9, 0.05)
set.seed(123987)
search <- foreach(i = a, .combine = rbind) %dopar% {
  cv <- cv.glmnet(mdlX, mdlY, family = "binomial", nfold = 10, type.measure = "deviance", parallel  = TRUE, alpha = i)
  data.frame(cvm = cv$cvm[cv$lambda == cv$lambda.1se], lambda.1se = cv$lambda.1se, alpha = i)
}
cv3 <- search[search$cvm == min(search$cvm), ]
md_dep <- glmnet(mdlX, mdlY, family = "binomial", lambda = cv3$lambda.1se, alpha = cv3$alpha)
#md_dep <- glmnet(mdlX, mdlY, family = "binomial", lambda = .24, alpha = .35)
coef(md_dep)

roc(newY, as.numeric(predict(md_dep, newX, type = "response")))


ROC_Final_EPDS_Com = roc(newY, as.numeric(predict(md_dep, newX, type = "response")))

ggroc(ROC_Final_EPDS_Com, colour = 'steelblue', legacy.axes = TRUE, size = 2) + 
  theme(text=element_text(size= 16)) +
  ggtitle(paste0('Elastic net regression with depression outcome ',
                 '(AUC = ', round(ROC_Final_EPDS_Com$auc,4),')'))

```

   
## EN Reg for ANX T1 only

```{r}
set.seed(123987)
n <- nrow(df_beck_pred_complete)
sample <- sample(seq(n), size = n * 0.5, replace = FALSE) 
train <- df_beck_pred_complete[sample,]
test <- df_beck_pred_complete[-sample,]
class(df_beck_pred_complete$pp_BECK)
# names(df_beck_pred_complete)
# any_na(df_beck_pred_complete)

mdlY <- (data.matrix(train[,"pp_BECK" ],))

mdlX <-scale(data.matrix(train[,c( "POB", "EDU", "SLEEP", "AGE_tot", "BMI_tot","PAR", "ILL", "DEPH", "SSRI", "FOD", "RS_14_tot", 
                                   "EPDS_32_tot", "BECK_32_tot"), ]))
newY <- (data.matrix(test[,"pp_BECK"], ))
newX <- scale(data.matrix(test[,c("POB", "EDU", "SLEEP", "AGE_tot", "BMI_tot","PAR", "ILL", "DEPH", "SSRI", "FOD", "RS_14_tot", 
                                  "EPDS_32_tot", "BECK_32_tot"), ]))



a <- seq(0.1, 0.9, 0.05)
set.seed(123987)
search <- foreach(i = a, .combine = rbind) %dopar% {
  cv <- cv.glmnet(mdlX, mdlY, family = "binomial", nfold = 10, type.measure = "deviance", parallel  = TRUE, alpha = i)
  data.frame(cvm = cv$cvm[cv$lambda == cv$lambda.1se], lambda.1se = cv$lambda.1se, alpha = i)
}
cv3 <- search[search$cvm == min(search$cvm), ]
#md_anx <- glmnet(mdlX, mdlY, family = "binomial", lambda = .15, alpha = .35)
md_anx <- glmnet(mdlX, mdlY, family = "binomial", lambda = cv3$lambda.1se, alpha = cv3$alpha)
coef(md_anx)
#class(newY)
roc(as.factor(newY), predict(md_anx, newX, type = "response", direction = ">"))
ROC_EN_BECK_Com = roc(as.factor(newY), predict(md_anx, newX, type = "response", direction = ">"))
as.factor(as.numeric(newY))

ggroc(ROC_EN_BECK_Com, colour = 'steelblue', legacy.axes = TRUE, size = 2) + 
  theme(text=element_text(size= 16)) +
  ggtitle(paste0('Elastic net regression with anxiety outcome ',
                 '(AUC = ', round(ROC_EN_BECK_Com$auc,4),')'))
```

### OUTPUT
```{r}
# 14 x 1 sparse Matrix of class "dgCMatrix"
#                     s0
# (Intercept) -1.9815367
# POB          .        
# EDU          .        
# SLEEP        .        
# AGE_tot      .        
# BMI_tot      .        
# PAR          .        
# ILL          .        
# DEPH         .        
# SSRI         .        
# FOD          .        
# RS_14_tot    .        
# EPDS_32_tot  .        
# BECK_32_tot  0.5191754


```



##  EN Reg for DEP HRV only

```{r}
set.seed(123987)  #123987 used in final model
n <- nrow(df_epds_pred_complete)
sample <- sample(seq(n), size = n * 0.8, replace = FALSE) 
train <- df_epds_pred_complete[ sample, ]
test <- df_epds_pred_complete[-sample,]
# str(train)
# summary(train)
# names(df_epds_pred_complete)
# any_na(df_epds_pred_complete)
# class(df_epds_pred_complete$pp_EPDS)


mdlY <- as.factor(data.matrix(train[,"pp_EPDS" ],))
mdlX <-scale (as.matrix(train[,c( "HRV_HFpow_1", "HRV_HFpow_2" , "HRV_LFpow_1" , "HRV_LFpow_2", "HRV_TP_1", "HRV_TP_2",
                                 "HRV_RMSSD_1", "HRV_RMSSD_2", "HRV_SDNN_1", "HRV_SDNN_2","HRV_LFHF_1", "HRV_LFHF_2"), ]))
newY <- as.factor(data.matrix(test[,"pp_EPDS"], ))
newX <- scale (as.matrix(test[,c("HRV_HFpow_1", "HRV_HFpow_2" , "HRV_LFpow_1" , "HRV_LFpow_2", "HRV_TP_1", "HRV_TP_2",
                                 "HRV_RMSSD_1", "HRV_RMSSD_2", "HRV_SDNN_1", "HRV_SDNN_2","HRV_LFHF_1", "HRV_LFHF_2"), ]))

### removed "RS_14_tot",
#### Elastic net 
a <- seq(0.1, 0.9, 0.05)
set.seed(123987)
search <- foreach(i = a, .combine = rbind) %dopar% {
  cv <- cv.glmnet(mdlX, mdlY, family = "binomial", nfold = 10, type.measure = "deviance", parallel  = TRUE, alpha = i)
  data.frame(cvm = cv$cvm[cv$lambda == cv$lambda.1se], lambda.1se = cv$lambda.1se, alpha = i)
}
cv3 <- search[search$cvm == min(search$cvm), ]
md_dep <- glmnet(mdlX, mdlY, family = "binomial", lambda = cv3$lambda.1se, alpha = cv3$alpha)
#md_dep <- glmnet(mdlX, mdlY, family = "binomial", lambda = .25, alpha = .25)
coef(md_dep)

roc(newY, as.numeric(predict(md_dep, newX, type = "response")))


ROC_Final_EPDS_Com = roc(newY, as.numeric(predict(md_dep, newX, type = "response")))

ggroc(ROC_Final_EPDS_Com, colour = 'steelblue', legacy.axes = TRUE, size = 2) + 
  theme(text=element_text(size= 16)) +
  ggtitle(paste0('Elastic net regression with depression outcome ',
                 '(AUC = ', round(ROC_Final_EPDS_Com$auc,4),')'))
```
### Output
```{r}
# 13 x 1 sparse Matrix of class "dgCMatrix"
#                      s0
# (Intercept) -1.05635786
# HRV_HFpow_1 -0.09625537
# HRV_HFpow_2  .         
# HRV_LFpow_1  .         
# HRV_LFpow_2  .         
# HRV_TP_1     .         
# HRV_TP_2     .         
# HRV_RMSSD_1  .         
# HRV_RMSSD_2  .         
# HRV_SDNN_1   .         
# HRV_SDNN_2   .         
# HRV_LFHF_1   .         
# HRV_LFHF_2   .  



# cv3$lambda.1se
# [1] 0.1430657
# > alpha = cv3$alpha
# > cv3$alpha
# [1] 0.75
```



## EN Reg for ANX HRV only

```{r}
set.seed(123987)
n <- nrow(df_beck_pred_complete)
sample <- sample(seq(n), size = n * 0.5, replace = FALSE) 
train <- df_beck_pred_complete[sample,]
test <- df_beck_pred_complete[-sample,]
class(df_beck_pred_complete$pp_BECK)
# names(df_beck_pred_complete)
# any_na(df_beck_pred_complete)

mdlY <- (data.matrix(train[,"pp_BECK" ],))

mdlX <-scale(data.matrix(train[,c( "HRV_HFpow_1", "HRV_HFpow_2" , "HRV_LFpow_1" , "HRV_LFpow_2", "HRV_TP_1", "HRV_TP_2",
                                   "HRV_RMSSD_1", "HRV_RMSSD_2", "HRV_SDNN_1", "HRV_SDNN_2","HRV_LFHF_1", "HRV_LFHF_2"), ]))
newY <- (data.matrix(test[,"pp_BECK"], ))
newX <- scale(data.matrix(test[,c("HRV_HFpow_1", "HRV_HFpow_2" , "HRV_LFpow_1" , "HRV_LFpow_2", "HRV_TP_1", "HRV_TP_2",
                                  "HRV_RMSSD_1", "HRV_RMSSD_2", "HRV_SDNN_1", "HRV_SDNN_2","HRV_LFHF_1", "HRV_LFHF_2"), ]))



a <- seq(0.1, 0.9, 0.05)
set.seed(123987)
search <- foreach(i = a, .combine = rbind) %dopar% {
  cv <- cv.glmnet(mdlX, mdlY, family = "binomial", nfold = 10, type.measure = "deviance", parallel  = TRUE, alpha = i)
  data.frame(cvm = cv$cvm[cv$lambda == cv$lambda.1se], lambda.1se = cv$lambda.1se, alpha = i)
}
cv3 <- search[search$cvm == min(search$cvm), ]
#md_anx <- glmnet(mdlX, mdlY, family = "binomial", lambda = .15, alpha = .35)
md_anx <- glmnet(mdlX, mdlY, family = "binomial", lambda = cv3$lambda.1se, alpha = cv3$alpha)
coef(md_anx)
#class(newY)
roc(as.factor(newY), predict(md_anx, newX, type = "response", direction = ">"))
ROC_EN_BECK_Com = roc(as.factor(newY), predict(md_anx, newX, type = "response", direction = ">"))
as.factor(as.numeric(newY))

ggroc(ROC_EN_BECK_Com, colour = 'steelblue', legacy.axes = TRUE, size = 2) + 
  theme(text=element_text(size= 16)) +
  ggtitle(paste0('Elastic net regression with anxiety outcome ',
                 '(AUC = ', round(ROC_EN_BECK_Com$auc,4),')'))
```

### OUTPUT
```{r}
#13 x 1 sparse Matrix of class "dgCMatrix"
                  # s0
# (Intercept) -1.882731
# HRV_HFpow_1  0.000000
# HRV_HFpow_2  .       
# HRV_LFpow_1  .       
# HRV_LFpow_2  .       
# HRV_TP_1     .       
# HRV_TP_2     .       
# HRV_RMSSD_1  .       
# HRV_RMSSD_2  .       
# HRV_SDNN_1   .       
# HRV_SDNN_2   .       
# HRV_LFHF_1   .       
# HRV_LFHF_2   .  
                   
# cv3$lambda.1se
# [1] 0.85
# > cv3$alpha
# [1] 0.15                  
```


### Group totals 
```{r}
table(df_epds_pred_complete$pp_EPDS)

table(df_beck_pred_complete$pp_BECK)

table(df_epds_pred_complete$EPDS_32_tot)
```


### Overlap between groups
```{r}
xtabs(~ppv6_EPDS_10_11R + ppv6_beck_score_R_01, data=BASIC_updated_22nov_na_rm)

xtabs(~ppv6_EPDS_10_11R + v32_EPDS_D_9R, data=df1)
```


### Calcualting percent change for DEP outcome. !! LOOK UP HRV_PerChg.Rmd or HRV_Table2.Rmd for updated analyses!!
```{r}
library(dplyr)
EPDS_HF_PC <- df_epds_pred_complete %>% 
  select(HRV_HFpow_1, HRV_HFpow_2) %>%
  mutate(HF_pct_chg = (100 * (HRV_HFpow_2 - (HRV_HFpow_1))/(HRV_HFpow_1)))


EPDS_LF_PC <- df_epds_pred_complete %>% 
  select(HRV_LFpow_1, HRV_LFpow_2) %>%
  mutate(LF_pct_chg = (100 * (HRV_LFpow_2 - (HRV_LFpow_1))/(HRV_LFpow_1)))


EPDS_TP_PC <- df_epds_pred_complete %>% 
  select(HRV_TP_1, HRV_TP_2) %>%
  mutate(TP_pct_chg = (100 * (HRV_TP_2 - (HRV_TP_1))/(HRV_TP_1)))

EPDS_LFHF_PC <- df_epds_pred_complete %>% 
  select(HRV_LFHF_1, HRV_LFHF_2) %>%
  mutate(LFHF_pct_chg = (100 * (HRV_LFHF_2 - (HRV_LFHF_1))/(HRV_LFHF_1)))

EPDS_RMSSD_PC <- df_epds_pred_complete %>% 
  select(HRV_RMSSD_1, HRV_RMSSD_2) %>%
  mutate(RMSSD_pct_chg = (100 * (HRV_RMSSD_2 - (HRV_RMSSD_1))/(HRV_RMSSD_1)))

EPDS_SDNN_PC <- df_epds_pred_complete %>% 
  select(HRV_SDNN_1, HRV_SDNN_2) %>%
  mutate(SDNN_pct_chg = (100 * (HRV_SDNN_2 - (HRV_SDNN_1))/(HRV_SDNN_1)))

```

```{r}

t.test(EPDS_HF_PC$HF_pct_chg  ~  df_epds_pred_complete$pp_EPDS)

t.test(EPDS_LF_PC$LF_pct_chg  ~  df_epds_pred_complete$pp_EPDS)

df_epds_pred_complete$pp_EPDS <- as.numeric(df_epds_pred_complete$pp_EPDS)

wilcox.test(EPDS_LFHF_PC$LFHF_pct_chg, df_epds_pred_complete$pp_EPDS)

t.test(EPDS_TP_PC$TP_pct_chg  ~  df_epds_pred_complete$pp_EPDS)

wilcox.test(EPDS_RMSSD_PC$RMSSD_pct_chg, df_epds_pred_complete$pp_EPDS)

wilcox.test(EPDS_SDNN_PC$SDNN_pct_chg, df_epds_pred_complete$pp_EPDS)
```

### Check histograms
```{r}
histogram(EPDS_HF_PC$HF_pct_chg)
histogram(EPDS_LF_PC$LF_pct_chg)
histogram(EPDS_LFHF_PC$LFHF_pct_chg)
histogram(EPDS_TP_PC$TP_pct_chg)
histogram(EPDS_RMSSD_PC$RMSSD_pct_chg)
histogram(EPDS_SDNN_PC$SDNN_pct_chg)


```



### Calcualting percent change for ANX outcome

```{r}
library(dplyr)
BECK_HF_PC <- df_beck_pred_complete %>% 
  select(HRV_HFpow_1, HRV_HFpow_2) %>%
  mutate(HF_pct_chg = (100 * (HRV_HFpow_2 - (HRV_HFpow_1))/(HRV_HFpow_1)))


BECK_LF_PC <- df_beck_pred_complete %>% 
  select(HRV_LFpow_1, HRV_LFpow_2) %>%
  mutate(LF_pct_chg = (100 * (HRV_LFpow_2 - (HRV_LFpow_1))/(HRV_LFpow_1)))


BECK_TP_PC <- df_beck_pred_complete %>% 
  select(HRV_TP_1, HRV_TP_2) %>%
  mutate(TP_pct_chg = (100 * (HRV_TP_2 - (HRV_TP_1))/(HRV_TP_1)))

BECK_LFHF_PC <- df_beck_pred_complete %>% 
  select(HRV_LFHF_1, HRV_LFHF_2) %>%
  mutate(LFHF_pct_chg = (100 * (HRV_LFHF_2 - (HRV_LFHF_1))/(HRV_LFHF_1)))

BECK_RMSSD_PC <- df_beck_pred_complete %>% 
  select(HRV_RMSSD_1, HRV_RMSSD_2) %>%
  mutate(RMSSD_pct_chg = (100 * (HRV_RMSSD_2 - (HRV_RMSSD_1))/(HRV_RMSSD_1)))

BECK_SDNN_PC <- df_beck_pred_complete %>% 
  select(HRV_SDNN_1, HRV_SDNN_2) %>%
  mutate(SDNN_pct_chg = (100 * (HRV_SDNN_2 - (HRV_SDNN_1))/(HRV_SDNN_1)))

```



```{r}
t.test(BECK_HF_PC$HF_pct_chg  ~  df_beck_pred_complete$pp_BECK)

t.test(BECK_LF_PC$LF_pct_chg  ~    df_beck_pred_complete$pp_BECK)

df_beck_pred_complete$pp_BECK <- as.numeric(df_beck_pred_complete$pp_BECK)
wilcox.test(BECK_LFHF_PC$LFHF_pct_chg,  df_beck_pred_complete$pp_BECK)

t.test(BECK_TP_PC$TP_pct_chg  ~   df_beck_pred_complete$pp_BECK)

wilcox.test(BECK_RMSSD_PC$RMSSD_pct_chg,  df_beck_pred_complete$pp_BECK)

wilcox.test(BECK_SDNN_PC$SDNN_pct_chg,  df_beck_pred_complete$pp_BECK)
```



### Check BECK histogram 
```{r}
histogram(df_beck_pred_complete$HRV_HFpow_1)
histogram(df_beck_pred_complete$HRV_HFpow_2)
histogram(df_beck_pred_complete$HRV_LFpow_1)
histogram(df_beck_pred_complete$HRV_LFpow_2)
histogram(df_beck_pred_complete$HRV_TP_1)
histogram(df_beck_pred_complete$HRV_TP_2)
histogram(BECK_HF_PC$HF_pct_chg)
histogram(BECK_LF_PC$LF_pct_chg)
histogram(BECK_LFHF_PC$LFHF_pct_chg)

df_beck_pred_complete$pp_BECK <- as.numeric(df_beck_pred_complete$pp_BECK)
wilcox.test(BECK_LFHF_PC$LFHF_pct_chg,  df_beck_pred_complete$pp_BECK)

histogram(BECK_TP_PC$TP_pct_chg)
histogram(BECK_RMSSD_PC$RMSSD_pct_chg)
wilcox.test(BECK_RMSSD_PC$RMSSD_pct_chg,  df_beck_pred_complete$pp_BECK)

histogram(BECK_SDNN_PC$SDNN_pct_chg)

wilcox.test(BECK_SDNN_PC$SDNN_pct_chg,  df_beck_pred_complete$pp_BECK)
```


